---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# {suntimes}

<!-- badges: start -->
<!-- badges: end -->

## What is it?

{suntimes} is an `R` API wrapper for solar event times from [sunrise-sunset.org](https://sunrise-sunset.org).  The returned variables are

* sunrise
* sunset
* solar noon
* day length
* civil twilight start
* civil twilight end
* nautical twilight start
* nautical twilight end
* astronomical twilight start
* astronomical twilight end

## Why was it made?

{suntimes} was made as a teaching aid to show students (i) how to interface with simple APIs, and (ii) how to author packages.

In practice, it is better to compute solar event times internally in `R` rather than rely on an external API.  Two recommended packages are

* [{suncalc}](https://github.com/datastorm-open/suncalc)
* [{oce}](https://github.com/dankelley/oce/)

## Usage limits and attribution

[sunrise-sunset.org](https://sunrise-sunset.org) provide their API free of charge on the understanding that you do not use the API in a manner that exceeds reasonable request volume, constitutes excessive or abusive usage.

Should you use the data provided by the API, you must show attribution with a link to https://sunrise-sunset.org.

## Installation

Install the development version from [GitHub](https://github.com/) by typing the following

```{r, eval=FALSE, echo=TRUE}
# Install devtools if needed
#install.packages("devtools")

# Install suntimes using devtools
devtools::install_github("afsarchowdhury/suntimes")
```

A [CRAN](https://cran.r-project.org/) version is unavailable at this time.

## Example

To return solar event times, the latitude and longitude of the location in question must be provided.  These can be obtained from most mapping platforms, such as [latlong.net](https://www.latlong.net/).  Alternatively, use [{tidygeocoder}](https://github.com/jessecambon/tidygeocoder/) to get these programmatically.

The latitude and longitude for Land's End is 50.065471 and -5.714856, respectively.  Using these values with no given date, solar event times for the current day is returned.

```{r example_01, eval=FALSE, echo=TRUE}
# Load suntimes
library(suntimes)

# Get times
suntimes(lat = 50.065471, lon = -5.714856)
```

If you supply a date, solar event times are returned for that date.  Note: date must be supplied as a string in YYYY-MM-DD format.

```{r example_02, eval=FALSE, echo=TRUE}
suntimes(lat = 50.065471, lon = -5.714856, date = "2023-08-31")
```

All times are returned in UTC.  For summer time adjustments, you must provide a timezone.  Note: `R` must recognise the timezone on your system.  Use `OlsonNames()` for valid timezones.

```{r example_03, eval=FALSE, echo=TRUE}
suntimes(lat = 50.065471, lon = -5.714856, date = "2023-08-31", timezone = "Europe/London")
```

Solar event times for multiple dates can be obtained using the function `suntimes_multiple()`.  Be mindful of the fair usage policy when using this function.

```{r example_04, eval=FALSE, echo=TRUE}
suntimes_multiple(
  lat = 50.065471,
  lon = -5.714856,
  dates = c("2023-08-31", "2023-09-01"),
  timezone = "Europe/London"
)
```

## Extracting time data

The data returned by {suntimes} is in POSIXct by default.  Sometimes, only the time data is needed.  The [{hms}](https://hms.tidyverse.org/index.html) package can be used to extract the time.

```{r example_05, eval=TRUE, echo=TRUE}
# Load package
library(suntimes)

# Get data
df <- suntimes(lat = 50.065471, lon = -5.714856, date = "2023-08-31", timezone = "Europe/London")

# Extract time only
hms::as_hms(df$results.solar_noon)
```

Similarly, to plot using [{ggplot2}](https://ggplot2.tidyverse.org/)

```{r example_06, eval=TRUE, echo=TRUE}
# Load package
library(ggplot2)

# Define sequence of dates
date_start <- as.Date("2023-01-01")
date_end <- as.Date("2023-12-31")
date_seq <- seq(date_start, date_end, 14)

# Get data
df <- suntimes_multiple(lat = 50.065471, lon = -5.714856, dates = date_seq, timezone = "Europe/London")

# Plot
ggplot(df, aes(x = date_seq, y = hms::as_hms(results.solar_noon))) +
  geom_line() +
  geom_point() +
  labs(x = "Date", y = "Solar noon")
```

## License

{suntimes} is released on a [GPLv3 license](https://www.gnu.org/licenses/gpl-3.0.en.html).

## Citation

```{r citation, eval=TRUE}
citation("suntimes")
```
